---
title: "Data Story 2"
author: "Sarra Hannachi"
date: "21/07/2021"
output: html_document
---


<style>
body {
text-align: justify}
title {
font-style: bold}
</style>

<p>&nbsp;</p>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

                ############## GDPR article violations' fines in Europe & UK ##############
## Description:  
The General Data Protection Regulation (EU) 2016/679 (GDPR) is a regulation in EU law on data protection and privacy in the European Union (EU) and the European Economic Area (EEA). It also addresses the transfer of personal data outside the EU and EEA areas. The GDPR aims primarily to give control to individuals over their personal data and to simplify the regulatory environment for international business by unifying the regulation within the EU.[1] Superseding the Data Protection Directive 95/46/EC, the regulation contains provisions and requirements related to the processing of personal data of individuals (formally called data subjects in the GDPR) who reside in the EEA, and applies to any enterprise—regardless of its location and the data subjects' citizenship or residence—that is processing the personal information of data subjects inside the EEA.
### Source: https://www.enforcementtracker.com/ 


## Data Cleaning

```{r, warning = FALSE, include=FALSE, message=FALSE}
## Basic data cleaning and organization ##

#import
library(readxl)
data <- read_excel("GDPR_data.xlsx")
#date
data$Date <- as.Date(data$Date)
#clean NA fines
data$Fines <- as.factor(data$Fines)
vec <- data$id[data$Fines == "Unknown" | data$Fines == "Only intention to issue fine"]
data <- data[data$id %in% vec == F,]
#Numerate fines
data$Fines <- as.numeric(gsub(",", "", data$Fines))
#Keep only 2020 & 2021
data <- data[1:556,-10]
View(data)
#type (10 unique)
data$Type <- as.factor(data$Type)
levels(data$Type)
#Authority (46 unique)
data$Authority <- as.factor(data$Authority)
levels(data$Authority)
#Country (31 unique)
data$Country <- as.factor(data$Country)
levels(data$Country)
#Controller
data$Controller <- as.factor(data$Controller)
levels(data$Controller)
t <- as.data.frame.table(table(data$Controller))
```

## PLAN:

```{r, warning = FALSE, include=FALSE, message=FALSE}
                        ### Highlights ###
#Viz 1#
# Fines in time series (2021)         => LINE CHART (Animated gif)

                        ### Story Body ###
#Viz 2#
# The most/least frequent violation types by the total GDPR fines amount
#Viz 3#
# countries/fines frequency?     => Progress BAR CHART  (Animated)
# Emphasize on the most frequent countries in top 10
#Viz 4#
# Make a column chart of fines distribution in % by Industry of Data Controller
```

## Highlights:

## Body:

### Graph 1: Transitioning through time

```{r}
#Data Transformation
library(dplyr)
by_day <- group_by(data[,c(4,5)], Date)
by_day <- summarise(by_day, TotalFines = sum(Fines))
#Fines data by YEAR
data_2020 <- by_day[1:150,]
data_2021 <- by_day[151:250,]
library(writexl)
write_xlsx(data_2021,"GDPR data 2021.xlsx")
data_2021$TotalFines <- data_2021$TotalFines/1000000
#Min Fine
data_2021$Date[data_2021$TotalFines == min(data_2021$TotalFines)]
#Max Fine
data_2021$Date[data_2021$TotalFines == max(data_2021$TotalFines)]
#Average fine by type
max(data$Fines[data$Type != "Insufficient legal basis for data processing" & data$Type != "Insufficient technical and organisational measures to ensure information security" & data$Type != "Non-compliance with general data processing principles" & data$Type != "Insufficient fulfilment of data subjects rights" & data$Type != "Insufficient fulfilment of information obligations"])
levels(data$Type)
```

```{r}
#load packages
library(gganimate)
library(ggplot2)
library(ggthemes)
library(extrafont)

# Graph 1: 
Sys.setlocale(category = "LC_ALL", locale = "english")
graph1 <- data_2021 %>%
  ggplot(aes(x=Date, y=TotalFines)) +
  geom_line(color="deeppink4" ,size = 1, alpha = 0.6)+
  labs(title = "Total Fines of GDPR Violations in Europe\n",
       subtitle = "Year: 2021\n",
       x = "\n\n\nDate",
       y = "Fines (Million Euros)\n\n",
       caption = "Source: GDPR Enforcement Tracker") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        panel.background = element_rect(fill = NA),
        plot.title = element_text(hjust=0,size=20))+
  scale_x_date(date_breaks="4 weeks", date_labels = "%b")
graph1

graph1.animated <- graph1 +
  transition_reveal(Date)

animate(graph1.animated, height = 400, width = 800, duration = 20, end_pause = 20)
anim_save("reveal.gif",animation = last_animation())
```

### Graph 2: Donut Chart of TYPES (%)

```{r}
#Dataframe for Fines by Type
by_t <- group_by(data[,c(9,5)], Type)
by_t <- summarise(by_t, Total_Fines = sum(Fines))
by_t <- mutate(by_t, Percentage = round(100*by_t$Total_Fines/sum(by_t$Total_Fines),2))
by_t <- arrange(by_t, desc=Percentage)
sum(by_t$Percentage)
```


### Graph 3: Racing Bar Charts (Country/weekly cumulative fines)

* 2020
```{r}
library(lubridate)
#original (2020)
df_2020 <- data[223:556,c(2,4,5)]
#daily quarter
df_2020$Quarter <- quarter(df_2020$Date)
#group by Quarter and Country
df_2020$Quarter [df_2020$Quarter == "1"] <- "Q1 2020"
df_2020$Quarter [df_2020$Quarter == "2"] <- "Q2 2020"
df_2020$Quarter [df_2020$Quarter == "3"] <- "Q3 2020"
df_2020$Quarter [df_2020$Quarter == "4"] <- "Q4 2020"
df_2020$Quarter <- as.factor(df_2020$Quarter)
#tidy data
df_2020$month <- month(df_2020$Date,label = F)
df_2020$Date <- paste("week", ceiling(day(df_2020$Date) / 7))
df_2020 <- transform(df_2020, week=paste(Date, month, "2020",sep=" "))
df_2020$week <- factor(df_2020$week, levels = unique(df_2020$week))
#transform data weekly
library(dplyr)
duff <- df_2020 %>% group_by(Country,week) %>% summarise(Total_fines=sum(Fines))
duff <- arrange(duff,week)
#Tidy data
library(reshape2)
tidy_ <- dcast(duff, Country ~ week, value.var = "Total_fines")
tidy_[is.na(tidy_)] <- 0
tidy_ <- rev(tidy_)
tidy_ <- tidy_[,c(57,1:56)]
td <- tidy_[,-1]
#save new dataframe
write.csv(tidy_,"tidy_2020.csv")
#cum sum 
td <- t(td)
for (j in 1:ncol(td)) {
   for (i in 2:nrow(td)) {
        td[i,j] <- td[i,j] + td[i-1,j]
    }
}
td <- t(td)
write.csv(td,"tidy_2020.csv")

```

* 2021
```{r}
#original (2021)
df_2021 <- data[1:222,c(2,4,5)]
df_2021$Date <- quarter(df_2021$Date)
#group by Quarter and Country
df_2021 <- df_2021 %>% group_by(Country, Date) %>% summarise(Total_fines=sum(Fines))
df_2021$Date <- as.character(df_2021$Date)
df_2021$Date [df_2021$Date == "1"] <- "Q1 2021"
df_2021$Date [df_2021$Date == "2"] <- "Q2 2021"
df_2021$Date [df_2021$Date == "3"] <- "Q3 2021"
df_2021$Date [df_2021$Date == "4"] <- "Q4 2021"
df_2021$Date <- as.factor(df_2021$Date)
```

## 2020 and 2021 combined together

```{r}
######♦ 2020 ########
library(lubridate)
#original (2020)
df_2020 <- data[223:556,c(2,4,5)]
#daily quarter
df_2020$Quarter <- quarter(df_2020$Date)
#group by Quarter and Country
df_2020$Quarter [df_2020$Quarter == "1"] <- "Q1 2020"
df_2020$Quarter [df_2020$Quarter == "2"] <- "Q2 2020"
df_2020$Quarter [df_2020$Quarter == "3"] <- "Q3 2020"
df_2020$Quarter [df_2020$Quarter == "4"] <- "Q4 2020"
df_2020$Quarter <- as.factor(df_2020$Quarter)
#tidy data
df_2020$month <- month(df_2020$Date,label = F)
df_2020$Date <- paste("week", ceiling(day(df_2020$Date) / 7))
df_2020 <- transform(df_2020, week=paste(Date, month, "2020",sep=" "))
df_2020$week <- factor(df_2020$week, levels = unique(df_2020$week))
######♦ 2021 ########
#original (2021)
df_2021 <- data[1:222,c(2,4,5)]
df_2021$Quarter <- quarter(df_2021$Date)
#group by Quarter and Country
df_2021$Quarter [df_2021$Quarter == "1"] <- "Q1 2021"
df_2021$Quarter [df_2021$Quarter == "2"] <- "Q2 2021"
df_2021$Quarter [df_2021$Quarter == "3"] <- "Q3 2021"
df_2021$Quarter [df_2021$Quarter == "4"] <- "Q4 2021"
df_2021$Quarter <- as.factor(df_2021$Quarter)
#tidy data
df_2021$month <- month(df_2021$Date,label = F)
df_2021$Date <- paste("week", ceiling(day(df_2021$Date) / 7))
df_2021 <- transform(df_2021, week=paste(Date, month, "2021",sep=" "))
df_2021$week <- factor(df_2021$week, levels = unique(df_2021$week))
#### merge #####
df_all <- rbind(df_2021,df_2020)
#transform data weekly
library(dplyr)
df_all <- df_all %>% group_by(Country,week) %>% summarise(Total_fines=sum(Fines))
df_all <- arrange(df_all,week)
View(df_all)
#Tidy data
library(reshape2)
tidy_all <- dcast(df_all, Country ~ week, value.var = "Total_fines")
tidy_all[is.na(tidy_all)] <- 0
tidy_all <- rev(tidy_all)
tidy_all <- tidy_all[,c(88,1:87)]
td <- tidy_all[,-1]
#save new dataframe
write.csv(tidy_all,"tidy_all.csv")
#cum sum 
td <- t(td)
for (j in 1:ncol(td)) {
   for (i in 2:nrow(td)) {
        td[i,j] <- td[i,j] + td[i-1,j]
    }
}
td <- t(td)
View(td)
countries <- as.data.frame(tidy_all$Country)
td <- cbind(countries,td)
colnames(td)[1] <- "Country"
write.csv(td,"tidy_2020_2021.csv")
```


### Graph 4: % Chart of top and least fined controllers' Industries

```{r include = FALSE}
#Dataframe for Fines by Controller
data$Industry <- as.character(data$Industry)
by_c <- group_by(data[,c(7,5)], Industry)
by_c <- summarise(by_c, Total_Fines = sum(Fines))
by_c <- mutate(by_c, Percentage = round(100*by_c$Total_Fines/sum(by_c$Total_Fines),2))
by_c <- arrange(by_c, desc=Percentage)
by_c$Industry[by_c$Industry == "Not assigned"] <- "Other"
by_c$Industry <- as.factor(by_c$Industry)
by_c$Total_Fines[11] /sum(by_c$Total_Fines)
```

* simple Pie Chart example:
```{r}
library(RColorBrewer)
library(plotrix)
pie3D(by_c$Percentage,
      col=brewer.pal(11,"Pastel2"),
      labels = by_c$Industry,
      labelcex = 1,
      explode=0.1,
      theta = 0.8,
      main="Percentage Distribution of Fines by Industry")
```
`












